---
title: "Exercise 11: Logistic Regression Solution"
editor: source
reference-location: margin
---

::: callout-note
Exercises are for practice purpose only.
:::



## Logistic Regression

<!-- ISL 4.6 -->

1. Suppose we collect data for a group of students in a statistics class with variables $X_1 =$hours studied, $X_2 =$undergrad GPA, and $Y =$receive an A. We fit a logistic regression and produce estimated coefficient, $\hat{\beta_0} = −6$, $\hat{\beta_1} = 0.05$, $\hat{\beta_2} = 1$. 

    (a) Estimate the probability that a student who studies for 40 hours and has an undergrad GPA of 3.5 gets an A in the class.
    
    (b) How many hours would the student in part (a) need to study to have a 50% chance of getting an A in the class?
    

```{r}
b0 <- -6
b1 <- 0.05
b2 <- 1
x2 <- 3.5
x1 <- 40
1 / (1 + exp(-(b0 + b1 * x1 + b2 * x2)))
```

The followings use `Boston` data set.
```{r}
library(ISLR2)
attach(Boston)
```

2. Create a new variable `crim01` having value 0 if `crim` is smaller than its median, and value 1 otherwise. Add the variable into the Boston data set.

```{r}
crim01 <- rep(0, length(crim))
crim01[crim > median(crim)] <- 1
Boston1 <- cbind(Boston, crim01)
```



3. Split the *new* `Boston` data set into training (~80%) and test sets (~20%).

```{r}
set.seed(1)
idx <- sample(nrow(Boston1), 406)
train <- Boston1[idx, ]
test <- Boston1[-idx, ]
```


3. Using the training set, fit logistic regression to predict whether a given census tract in the test set has a crime rate above or below the median.

```{r}
fit <- glm(crim01 ~ . - crim01 - crim, data = train, 
           family = binomial)
prob <- predict(fit, test, type = "response")
pred <- 1 * (prob > 0.5)
table(pred, test$crim01)
mean(pred == test$crim01)
```

