---
title: "Exercise 4: Diagnostics Solution"
editor: source
reference-location: margin
---

::: callout-note
Exercises are for practice purpose only.
:::



## Diagnostics

<!-- Yu -->
```{r}
library(ISLR2)
```

1. This problem involves the `Boston` data set. We now try to predict per capita crime rate using the median value of owner-occupied homes.

a. Fit a simple linear regression model. Check if there are any leverage points and outliers.

```{r}
lmout <- lm(crim ~ medv, data = Boston)
hat_i <- hatvalues(lmout)
p <- 2
n <- dim(Boston)[1]
hat_i[hat_i > 2 * p/n]
r_student <- rstudent(lmout)
r_student[r_student > 2]
```

b. Check the normality, constant variance and linearity.

```{r}
## normality
car::qqPlot(lmout, id = TRUE, col.lines = "blue", 
            reps = 1000, ylab = "Ordered R-Student Residuals", pch = 16)
```


```{r}
## normality
rstud <- rstudent(lmout)
car::densityPlot(rstud)
```


```{r}
## constant variance
car::scatterplot(fitted(lmout), rstudent(lmout), 
            smooth = list(span = 2/3, lwd.smooth = 3, lwd.spread = 4), 
            regLine = FALSE, boxplots = FALSE, main = "crim as y",
            xlab = "Fitted Values", ylab = "R-Student Residuals")
abline(a = 0, b = 0, col = 2, lwd = 2)
```

```{r}
## linearity
car::crPlots(lmout, ylab = "partial residual", grid = FALSE, main = "")
```


c. Transformming $y$ to satisfy linear regression assumptions.




```{r}
lmout_log <- lm(log(crim) ~ poly(medv, degree = 2, raw = TRUE), data = Boston)
```

```{r}
car::densityPlot(rstudent(lmout_log))
```

```{r}
car::scatterplot(fitted(lmout_log), rstudent(lmout_log), 
            smooth = list(span = 2/3, lwd.smooth = 3, lwd.spread = 4), 
            regLine = FALSE, boxplots = FALSE, main = "log(crim) as y, medv^2 as x",
            xlab = "Fitted Values", ylab = "R-Student Residuals")
abline(a = 0, b = 0, col = 2, lwd = 2)
```

```{r}
car::crPlots(lmout_log, ylab = "partial residual", grid = FALSE, main = "")
```


