---
title: "Exercise 5: Bootstrapping Solution"
editor: source
reference-location: margin
---

::: callout-note
Exercises are for practice purpose only.
:::



## Bootstrapping Standard Deviation

<!-- ISL 5.9 -->


1. Based on the `Boston` data set, provide an estimate for the population
mean of `medv`. Call the estimate $\hat{\mu}$.

```{r}
library(ISLR2)
attach(Boston)
(mu_hat <- mean(medv))
```


2. Provide an estimate of the standard error of $\hat{\mu}$. Interpret this
result. *Hint: We can compute the standard error of the sample mean by dividing the sample standard deviation by the square root of the number of observations.*

```{r}
(se <- sd(medv)/sqrt(length(medv)))
```


3. Estimate the standard error of $\hat{\mu}$ using the bootstrap. 

<!-- *Hint: We can use [boot::boot()](https://rdrr.io/cran/boot/man/boot.html) function.* -->

```{r}
boot.fn <- function(data, idx) mean(data[idx])
library(boot)
bstrap <- boot::boot(data = medv, statistic = boot.fn, R = 1000)
bstrap
## bias = bt_mean - obs_mean
```

```{r}
#| message: false
library(tidymodels)
boot_sample <- Boston |> 
    # specify the variable of interest
    specify(response = medv) |>  
    # generate 15000 bootstrap samples
    generate(reps = 1000, type = "bootstrap") |>  
    # calculate the median of each bootstrap sample
    calculate(stat = "mean")
sd(boot_sample$stat)
```

