---
title: "Exercise 6: Polynomial Regression Solution"
editor: source
reference-location: margin
---

::: callout-note
Exercises are for practice purpose only.
:::



## Polynomial Regression


<!-- LRA 7.1 -->
1. Consider the values of x shown below: 
$$x = 1.00, 1.70, 1.25, 1.20, 1.45, 1.85, 1.60, 1.50, 1.95, 2.00$$ Suppose that we wish to fit a second-order model using these levels for the regressor variable $x$. Calculate the correlation between $x$ and $x^2$. Do you see any potential difficulties in fitting the model?


```{r}
x <- c(1.00, 1.70, 1.25, 1.20, 1.45, 1.85, 1.60, 1.50, 1.95, 2.00)
x2 <- x ^ 2
cor(x, x2)
```


<!-- ISL 5.15 -->
2. Is there evidence of non-linear association between any of the predictors and the response `crim` in the `Boston` data set? To answer this question, for each predictor $X$, fit a model of the form 
$$Y = \beta_0 + \beta_1X+\beta_2X^2+\beta_3X^3+\epsilon$$


```{r}
library(ISLR2)
name <- names(Boston)
# 'degree' must be less than number of unique points
for (i in 2:13) {
  if (name[i] == "chas") next
  print(name[i])
  poly_reg <- lm(Boston$crim ~ poly(Boston[[i]], degree = 3))
  print(summary(poly_reg), digits = 2)
}
```

