---
title: "Exercise 8: Categorical Variables Solution"
editor: source
reference-location: margin
---

::: callout-note
Exercises are for practice purpose only.
:::

<!-- ## Conceptual -->
<!-- ISL 3.3 -->

Suppose we have a data set with five predictors, $X_1 = \text{GPA}$, $X_2 = \text{IQ}$, $X_3 = \text{Level}$ (1 for College and 0 for High School), $X_4 =$ Interaction between GPA and IQ, and $X_5 =$ Interaction between GPA and Level. The response is starting salary after graduation (in thousands of dollars). Suppose we use least squares to fit the model, and get
$b_0 = 50$, $b_1 = 20$, $b_2 = 0.07$, $b_3 = 35$, $b_4 = 0.01$, $b_5 = −10.$.

a. Which answer is correct, and why?

    i. For a fixed value of IQ and GPA, high school graduates earn more, on average, than college graduates.
  
    ii. For a fixed value of IQ and GPA, college graduates earn more, on average, than high school graduates.
  
    iii. For a fixed value of IQ and GPA, high school graduates earn more, on average, than college graduates provided that the GPA is high enough.
    
    iv. For a fixed value of IQ and GPA, college graduates earn more, on average, than high school graduates provided that the GPA is high enough.
    

```{r}
## iii
```

b. Predict the salary of a college graduate with IQ of 110 and a GPA of 4.0.

```{r}
gpa <- 4
iq <- 110
level <- 1
50 + 20 * gpa + 0.07 * iq + 35 * level + 0.01 * (gpa * iq) - 10 * (gpa * level)
```


c. True or false: Since the coefficient for the GPA/IQ interaction term is very small, there is very little evidence of an interaction effect. Justify your answer.

```{r}
## FALSE
## 1. We must examine the p-value of the regression coefficient to
## determine if the interaction term is statistically significant or not.
## 2. The range of IQ is always between 100~200, which is wider range than GPA(0~4).
## Therefore, although the coefficient is small, with GPA fixed, the interaction term can have
## significant effects.
```